\noindent \textbf{Prior NSF Support.} PI Cong Liu  has recently received the following NSF grant:

\begin{itemize}
\vspace{-2mm}
\item \textit{CSR: Small: Predictable Real-Time Computing in GPU-enabled Systems}, CNS-1527727, 10/15-9/18.
\vspace{-2mm}
\end{itemize}

This proposal addresses the challenge of applying GPUs in safety-critical systems, which ensures predictable real-time correctness. This goal will be achieved by (i) establishing GPU-aware resource allocation methods that yield quantifiable guarantees on real-time correctness, and (ii) implementing an ecosystem of GPU resource management that enables GPUs to be predictably utilized in a real-time multi-tasking environment. This grant will be used to support two PhD students (including a female student) for three years. 

?Hank: I think it suffices to cite one project that is most related to the proposed research, as space might be an issue.?

\vspace{0.05cm}

\noindent{\bf XPS: FULL: CCA: Collaborative Research: CASH: Cost-aware Adaptation of Software and Hardware} (CCF 1439156, PI: Hoffmann, 09/2014-08/2018, \$300,000)\\
\noindent{\bf EAGER: HAWKEYE}: A Cross-Layer Resilient Architecture to Tradeoff Program Accuracy and Resilience Overheads <HANK: Fill in details>\\
\noindent{\bf CNS: SMALL: BreezeFS} \\
\noindent{\bf II-New: RIVER: A Research Infrastructure to Explore Volatility, Energy-Efficiency, and Resilience} (CNS 1405959, PI: Chien, Co-PIs: Foster, Gunawi, Hoffmann, Scott, 07/2014-07/2017, \$997,432) 

All projects advance the study of adaptive computing systems.  CASH
studies the modeling and allocation of fine-grain hardware structures
to produce cost savings in infrastructure as a service (IaaS) clouds.
HAWKEYE studies the ability to dynamically trade program accuracy for
error detection overhead.  BreezeFS studies adaptive management of
multi-store systems for clouds.  Among other things, RIVER's
infrastructure supports study of large-scale adaptation to meet power
and performance goals in distributed computing systems.  Both projects
began less than 6 months ago.

\noindent {\em Intellectual Merit:} In the CASH project we have
developed novel hardware structures for performance monitoring as well
as novel combinations of machine learning and control theory for
allocating fine-grain hardware resources and minimizing costs.  Both
HAWKEYE and BreezeFS are in their early stages, but both will study
application of adaptive tradeoff management to new problems
(resilience and storage management, respectivel).  

% --------------
\noindent {\em Broader Impact:} The CASH project has already resulted
in four pulbished papers from the University of Chicago
\cite{kim-cpsna,POET,FSE2015,JouleGuard} and one submitted publication
whose outcome is still pending as of this writing.