%%%%%%%%% MASTER -- compiles the 4 sections

\documentclass[10pt,letterpaper]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pagestyle{plain}                                                      %%
%%%%%%%%%% EXACT 1in MARGINS %%%%%%%                                   %%
\setlength{\textwidth}{6.5in}     %%                                   %%
\setlength{\oddsidemargin}{0in}   %% (It is recommended that you       %%
\setlength{\evensidemargin}{0in}  %%  not change these parameters,     %%
\setlength{\textheight}{8.5in}    %%  at the risk of having your       %%
\setlength{\topmargin}{0in}       %%  proposal dismissed on the basis  %%
\setlength{\headheight}{-40pt}      %%  of incorrect formatting!!!)      %%
\setlength{\headsep}{65pt}         %%                                   %%
\setlength{\footskip}{0.3in}       %%                                   %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                   %%
\newcommand{\required}[1]{\section*{\hfil #1\hfil}}                    %%
\renewcommand{\refname}{\hfil References Cited\hfil}                   %%
\bibliographystyle{plain}                                              %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%PUT YOUR MACROS HERE

\usepackage{hyphenat}
\usepackage{layout}

\usepackage{verbatim} 
\usepackage{cite}
\usepackage[pdftex]{graphicx}
\usepackage{setspace}
\usepackage[cmex10]{amsmath}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{algorithmic}
\usepackage{algorithm}

\usepackage{subfig}
\usepackage{caption}

%\usepackage[caption=false]{subfig}

%\usepackage{subcaption}
%\captionsetup{compatibility=false}

\usepackage{mathpazo}

\linespread{1.05}      % Palatino needs more leading (space between lines)

\usepackage{euler}

\usepackage{url}

%\usepackage{courier}
%\renewcommand{\familydefault}{\sfdefault}

%\usepackage[compact]{titlesec}
%\addtolength{\itemsep}{-0.05in}
%\usepackage{subcaption}
\usepackage{paralist}

\usepackage{tikz}

\usepackage{pgfplots}
% options for pgfplots
\pgfplotsset{compat=1.8,compat/show suggested version=false}
\usetikzlibrary{calc,trees,arrows,patterns,plotmarks,shapes,snakes,er,3d,automata,backgrounds,topaths,decorations.pathmorphing,decorations.markings}
%\pgfplotsset{compat=newest}
\pgfplotsset{
   /pgfplots/bar  cycle  list/.style={/pgfplots/cycle  list={%
        {black,fill=black!30!white,mark=none},%
        {black,fill=red!30!white,mark=none},%
        {black,fill=green!30!white,mark=none},%
        {black,fill=yellow!30!white,mark=none},%
        {black,fill=brown!30!white,mark=none},%
     }
   },
}
\usetikzlibrary{external}
\tikzexternalize[prefix=out/]
\tikzexternalize
% don't externalize todonotes
%\makeatletter
%\renewcommand{\todo}[2][]{\tikzexternaldisable\@todo[#1]{#2}\tikzexternalenable}
%\makeatother
% end of externalization
\usetikzlibrary{patterns}
\usepgfplotslibrary{groupplots}
\pgfplotsset{
every axis label/.append style={font=\footnotesize},
tick label style={font=\footnotesize},
}
%\usepackage[normalem]{ulem}
%\setlength{\abovecaptionskip}{2pt plus 2pt minus 2pt}
%\setlist{noitemsep,topsep=0pt}

\usepackage{wrapfig}

%\usepackage[utf8]{inputenc}


%---------------------------------------------------------------------------------%
%*********************************************************************************%
%---------------------------------------------------------------------------------%


\title{\textbf{CSR:Small:Exploring Timing-Energy Tradeoffs on Heterogeneous Computing Platforms}}


\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
\let\endchangemargin=\endlist

\hyphenation{op-tical net-works semi-conduc-tor}

%\pagestyle{fancy}
%\includeonly{NSFsumm}

\date{\vspace{4mm}}

%\doublespace

\begin{document}

%\layout

\maketitle
\pagenumbering{arabic}


%---------------------------------------------------------------------------------%
%*********************************************************************************%
%---------------------------------------------------------------------------------%
\newpage 

\begin{center}
\Large{\textbf{Project Summary}}
\end{center}

\paragraph{Summary}
The physical constraints of thermal and power dissipation have led to
the emergence of \emph{dark silicon} in embedded processor designs.
Dark silicon refers to the fact that not all transistors can be
powered at full speed all the time --- doing so would destroy the chip
--- so some set of transistors must be kept dark --- unpowered.  To
address this issue, embedded processor designers have turned to
heterogeneous processor designs which combine powerful, but energy
inefficient cores, with slow, but energy efficient ones.

The combination of dark silicon and heterogeneous processor designs
creates both a challenge and an opportunity for embedded systems
designers and embedded operating systems in particular.  Embedded
operating systems must support hard real-time requirements and work to
reduce energy consumption.  The challenge of heterogeneous designs is
that real time-schedulers must now incorporate knowledge of differing
core types and reason about the diverse design space of scheduling
options that arise from multiple core types on the same chip.  The
opportunity is that dark silicon effectively creates a ``reserve'' of
additional computing power that can be brought to bear in short bursts
(but is not sustainable without damaging the hardware).  

\emph{This proposal's goal is to create new real-time scheduling
  algorithms for embedded systems that both address the challenge and
  exploit the opportunity of heterogeneous processor designs and dark
  silicon.}


\paragraph{Intellectual Merit} 
We seek to develop new real-time scheduling algorithms for
heterogeneous multicores that (1) are more energy efficient than the
current state-of-the-art and (2) can provide a new capability:
guaranteeing both timing and energy consumption for embedded
workloads.  While prior work has addressed real-time support on
heterogeneous architectures, it has primarily focused on porting
existing techniques developed for homogeneous systems and ensuring
that they can still meet hard real-time guarantees.  When ported to
heterogeneous systems, existing techniques are more energy efficient,
but this gain is entirely due to the processor architecture and not an
inherent feature of the algorithm.  The key insight in our proposed
work will be to start from scratch and incorporate knowledge of
heterogeneity and energy into our proposed algorithms from the
beginning of the project.  This approach promises to lead to new
algorithms that have better energy efficiency and offer new
capabilities.

We plan to evaluate these algorithms both analytically and
empirically.  Our empirical test case will be automotive computing, a
rich embedded processing domain that includes safety critical tasks,
like collision detection, and best effort tasks like infotainment.  We
will test the energy savings and schedulability of our algorithms by
implementing them for an state-of-the-art ARM big.LITTLE heterogeneous
embedded multicore and using them to schedule a representative
automotive workload.  We will demonstrate the proposed algorithms'
energy efficiency and ability to increase schedulability for the
automotive workload on real heterogeneous processors.

\paragraph{Broader Impacts}
The proposed research agenda provides for significant broader impacts
related to (1) curriculum development and student training through
integration of modules in existing real-time, energy-aware, and system
design courses, and (2) outreach through established enrichment
programs at the participating universities. The proposed research
benefits society by directly addressing the needs of smart
cyberphysical systems such as self-driving cars, embedded medical
devices, and intelligent robotics.  Such systems play critical roles
in manufacturing, defense, and healthcare. The proposed research
benefits cyberphysical systems by (1) enabling greater functionality
and new capabilities for the same cost and (2) greatly reducing system
energy consumption, perhaps leading to medical devices that last
longer or systems that need to be replaced less often.

\paragraph{Keywords:} 
Real-time; operating systems; embedded; energy; heterogeneous multicore; scheduling

%---------------------------------------------------------------------------------%
%*********************************************************************************%
%---------------------------------------------------------------------------------%
\newpage 

\section{Introduction}
\label{sec:intro}
\input{introduction.tex}

%*********************************************************************************%
%*********************************************************************************%

\section{Background and Related Work}
\label{sec:motivation}
\input{background.tex}

%*********************************************************************************%

\subsection{Hardware Platform}
\label{sec:hardware}
\input{hardware.tex}

%*********************************************************************************%

\subsection{Real-Time Scheduling Algorithms and Schedulability Tests}
\label{sec:model}
\input{model.tex}


%*********************************************************************************%

\subsection{Related Work}
\label{sec:relatedwork}
\input{relatedwork.tex}

%*********************************************************************************%
%*********************************************************************************%

\section{Detailed Research Plan}
\label{sec:researchplan}
\input{researchplan.tex}

%*********************************************************************************%

\subsection{Step 1: Identifying Energy-Efficient Configurations that Guarantee Timing}
\label{sec:step1}
\input{step1.tex}

%*********************************************************************************%

\subsection{Step 2: Energy Optimization and Guarantees}
\label{sec:step2}
\input{step2.tex}

%*********************************************************************************%

\subsection{Step 3: Implementation and Overhead-Conscious Schedulability Studies}
\label{sec:step3}
\input{step3.tex}

%*********************************************************************************%

\subsection{Step 4: Conduct Case Studies}
\label{sec:step4}
\input{step4.tex}

%*********************************************************************************%

\subsection{Timeline}
\label{sec:timeline}
\input{timeline.tex}

%*********************************************************************************%
%*********************************************************************************%

\section{Broader Impact of the Proposed Work}
\label{sec:impact}
\input{impact.tex}


%*********************************************************************************%
%*********************************************************************************%

\section{Results From Previous and Current NSF Support}
\label{sec:prior} 
\input{prior.tex}

%*********************************************************************************%
%*********************************************************************************%
\newpage
\pagenumbering{gobble}

\bibliographystyle{plain} 
\bibliography{b1,b2,b3,b4}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\pagenumbering{gobble}
\thispagestyle{empty}

\required{Cong Liu}
\vspace{-3mm}
\begin{center}
Department of Computer Science, University of Texas at Dallas, Richardson, TX 75081
http://www.utdallas.edu/$\sim$cong/
\end{center}


\noindent \textbf{Professional Preparation}\

	\begin{tabular}{l l l}
		Wuhan Univ. of Technology &  Computer Science & B.S. with honor, 2005 \\
		Auburn Univ. & Computer Science & M.S., 2008 \\
		Univ. of North Carolina at Chapel Hill & Computer Science & Ph.D., 2013\\ \\
	\end{tabular}

\noindent \textbf{Appointments}\

	\hspace{1mm} Assistant Professor, Department of Computer Science, University of Texas at Dallas, 9/13-present. \\
	
\noindent \textbf{Products} 

\hspace{-4mm} \textbf{(i) Most Relevant Publications}\  
	\vspace{-1mm}
	\begin{itemize}

\item 
Husheng Zhou and Cong Liu, ``\textit{Task Mapping in Heterogeneous Embedded Systems for Fast Completion Time}'', Proceedings of the 14th ACM International Conference on Embedded Software (\underline{EMSOFT}), 2014.

\item
Guangmo Tong and Cong Liu, ``\textit{Supporting Soft Real-Time Sporadic Task Systems on Heterogeneous Multiprocessors with No Uilitzation Loss}'', IEEE Transactions on Parallel and Distributed Systems (\underline{TPDS}), 2015.

	
\item
Husheng Zhou, Guangmo Tong, and Cong Liu, ``\textit{GPES: A Preemptive Execution System for GPGPU Computing}'', Proceedings of the 21st IEEE Real-Time and Embedded Technology and Applications Symposium (\underline{RTAS}), 2015.


\item  	Cong Liu, Jian Li, Wei Huang, Juan Rubio, and Evan Speight, ``\textit{Power-Efficient Time-Sensitive Mapping in CPU/GPU Heterogeneous Systems},'' Proceedings of the 21st International Conference on Parallel Architectures and Compilation Techniques (\underline{PACT}), pp. 23-32, 2012.

\item Guangmo Tong and Cong Liu, ``\textit{Supporting Read/Write Applications in Embedded Real-time Systems via Suspension-aware Analysis}'', Proceedings of the 14th ACM International Conference on Embedded Software (\underline{EMSOFT}), 2014.





\end{itemize}
\hspace{-4mm} \textbf{(ii) Additional Products}\  
\vspace{-1mm}
	\begin{itemize}
	


%\vspace{-2mm}
%\item 	Liang He, Linghe Kong, Siyu Lin, Shaodong Ying, Yu Gu, Tian he, and Cong Liu, ``\textit{Reconfiguration-Assisted Charging in Large-Scale Lithium-ion Battery Systems},'' Proceedings of the 5th ACM/IEEE International Conference on Cyber-Physical Systems \sloppy(\underline{ICCPS}), 2014.


\item
Jianjia Chen, Wenhung Huang, and Cong Liu, ``\textit{K2U: A General Framework from k-Point Effective Schedulabiliy Analysis to Utilization-based Tests}'', Proceedings of the 36th IEEE Real-Time Systems Symposium (\underline{RTSS}), 2015

\item
Jianjia Chen and Cong Liu, ``\textit{Fixed-Relative-Deadline Scheduling of Hard Real-Time Tasks with Self-Suspen- sions}'', Proceedings of the 35th IEEE Real-Time Systems Symposium (\underline{RTSS}), 2014.


\item 
Cong Liu and Jianjia Chen, ``\textit{Bursty-Interference Analysis Techniques for Analyzing Complex Real-Time Task Models}'', Proceedings of the 35th IEEE Real-Time Systems Symposium (\underline{RTSS}), 2014.

\vspace{-2mm}
\item 	Cong Liu and James Anderson, ``An O(m) Analysis Technique for Supporting Real-Time Self-Suspending Task Systems,'' Proceedings of the 33th IEEE Real-Time Systems Symposium (\underline{RTSS}), pp. 373-382, 2012.

%\vspace{-2mm}
%\item 	Cong Liu and James Anderson, ``\textit{Supporting Soft Real-Time DAG-based Systems on Multiprocessors with No Utilization Loss}," Proceedings of the 31st IEEE Real-Time Systems Symposium (\underline{RTSS}), pp. 3-13, 2010.

\vspace{-2mm}
\item Cong Liu and James Anderson, ``Task Scheduling with Self-Suspensions in Soft Real-Time Multiprocessor Systems,'' Proceedings of the 30th IEEE Real-Time Systems Symposium (\underline{RTSS}), pp.425-436, 2009,  \textbf{Best Student Paper Award.}

\end{itemize}
	
\newpage 

\noindent \textbf{Synergistic Activities}\
	\begin{changemargin}{4mm}{2mm}
	\textbf{Service to the Scientific and Engineering Community}\
	\vspace{-1mm}
		\begin{changemargin}{2mm}{2mm}	
		-  TPC, IEEE Real-Time and Embedded Technology and Applications Symposium (RTAS), 2015, 2016. \\
		 -   TPC, IEEE Real-Time Systems Symposium (RTSS), 2014, 2015. \\
			- TPC, Euromicro Conference on Real-Time Systems  (ECRTS), 2016.
			% Technical Program Committee Member, International Conference on High Performance Compilation, Computing and Communications, 2014.
			
			%\textit{Reviewer}: IEEE Transactions on Computers, IEEE Transactions on Parallel and Distributed Systems; IEEE Transactions on Industrial Informatics; ACM Transactions on Embedded Computing Systems; Real-Time Systems; Journal of Systems Architecture; Journal of Parallel and Distributed Computing; IEEE Real-TIme and Embedded Technology and Applications Symposium; IEEE Real-Time Systems Symposium; IEEE Real-TIme and Embedded Computing Systems and Applications; Euromicro Conference on Real-Time Systems; IEEE International Conference on Computer Communications;
			
		\end{changemargin}
	\end{changemargin}
	
	
\noindent \textbf{Collaborators and Other Affiliations}\\

\vspace{-2mm}
	\hspace{-2mm} \textbf{(i) Collaborators}\  
	\begin{changemargin}{8mm}{2mm}
	\vspace{-2mm}
	Dr. Evan Speight, IBM Research Austin Lab; \\
	Dr. Jian Li IBM Research Austin Lab;\\
	 Dr. Tian He, University of Minnesota, Twin City; \\
	  Dr. Jianjia Chen, TU Dortmund (German);\\
	   Dr. Yu Gu, Singapore University of Technology and Design (Singapore);\\
	    Dr. Wei Gao, University of Tennessee, Knoxville; \\
	    Dr. Shinpei Kato, Nagoya University (Japan).
	\end{changemargin}
	
	\hspace{-4mm} \textbf{(ii) Graduate and Post-doctoral Advisors}\  
	\begin{changemargin}{8mm}{2mm}
	Dr. James Anderson, Professor, Department of Computer Science, University of North Carolina at Chapel Hill.
	\end{changemargin}
	
	\hspace{-4mm} \textbf{(iii) Thesis Advisor and Post-graduate Scholar Sponsor}\  
	\begin{changemargin}{8mm}{2mm}

	Mr. Guangmo Tong, PhD Student, Computer Science, UTD (since 2014)
		
	Mr. Husheng Zhou, PhD Student, Computer Science, UTD (since 2014)	
	
	Ms. Xia Zhang, PhD Student, Computer Science, UTD (since 2014)	
	
	Mr. Yuchuan Liu, PhD Student, Computer Science, UTD (since 2015)

	Mr. Gbadebo Ayoade, Masters Student, Computer Science, UTD (since 2013)	

	\end{changemargin}


%---------------------------------------------------------------------------------%
%---------------------------------------------------------------------------------%
\newpage
\pagenumbering{gobble}

\begin{center}
\Large{\textbf{\\Facilities, Equipment, and Other Resources}}
\end{center}

\paragraph{General computing environment.} The Department of Computer Science hosts more than 40 research labs with access 24 hours a day, 7 days a week; 2 open labs for graduate students; and a 128-seat open access lab for undergraduate students. These labs consist of over 500 state-of-the-art, high-performance workstations and high-end PCs, all connected via gigabit Ethernet switches with a redundant fiber up-link to provide fast access to the campus network and the Internet. Nine classrooms and one large lecture hall with the latest computer and audio-visual equipment are available. Academic coursework, project, and computing systems are comprised of Linux x86\_64, Windows Server, and Solaris 10 executing on a collection of physical servers and virtual server private clouds. The servers are connected via fiber channel and iSCSI to a thin-provisioned 14TB 3PAR mesh-active storage array. 

\paragraph{Specific computing facilities.} Computer Science Open Access Lab: This lab is accessible to all students currently enrolled in any course in the department of computer science (128 seats). The lab consists of 128 Dell OptiPlex 980 machines with i7-870 Processor (8M Cache, 2.93 GHz), 4GB RAM, 22'' (Dell P2210) flat-panel display, Windows 7, 64-bit; Two HP Laserjet 9040 printers are also located in the lab. The lab is open from 8am to 11pm, M-F, and 11pm-7pm Sat/Sun. The Lab is maintained by the CS Tech Staff. A monitor (typically a graduate student hired half-time) is on premises all the time.

Graduate Windows Lab: This 15 seat lab is accessible to all graduate students.  Equipment consists of Dell Precision T1650 computers with Intel XEON CPU E3 1225 v2, @3.20GHz, 8GB RAM, 24'' flat-panel monitors, Windows 7 Enterprise and an HP Laserjet P4015n printer. The lab is accessible 24/7 with smart card access.

Project Design Lab: This 20 seat lab is accessible to all CS undergraduate/graduate students who take senior design project class. It consists of  Dell Optiplex 760 machines with Intel Core2 Duo, E8400@3.00GHz, 2GB RAM, 17''flat-panel monitors, Windows 7 Enterprise. The lab is accessible 24/7 with smart card access.

General Access Systems: The Department also maintains several systems in its server room to which remote access is provided. These include the Network Programming System that are ccessible to all CS students enrolled in network programming and parallel processing courses (45 systems); Students access these systems via SSH; All these 45 systems run Linux CentOS 6.2 x86\_64. The Department also maintains two large Linux compute servers (Linux CentOS 6.2 x86\_64, Dell PowerEdge R710 w/2 six-core 2.53GHz Intel Xeon, 32GB RAM) and two large Solaris computer servers (Sun Fire V440 w/4x1GHz UltraSPARC-IIIi, 8GB RAM). 

Server clouds: The Department also maintains a general purpose server cloud, a VMware vSphere 4 operating on 5 ESX hosts with dual 6-core processor 128GB RAM servers, used for teaching the cloud computing related courses as well as a network security courses server cloud,  a VMware vSphere 4 executing on 4 processors of 6-core Opterons, 128GB RAM, and four 4Gb redundant fiber channel SAN connections.

In addition to the above labs and facilities, CS students also have access the UTDesign Studio that is equipped with modern computing facilities and equipment with more than 30,000 square feet of dedicated space, where students and corporate partners can create, innovate, design, build and learn. Further information can be found at $http://www.utdallas.edu/utdesign/corporate/studio/$. The UTDesign studio is primarily used by undergraduate seniors taking the capstone project course.

A 30' x 20' lab is available for Dr. Liu's students. The lab can comfortably accommodate 8 students and the computing equipment. Eight workstations equipped with state-of-the-art NVIDIA GPUs currently in the PI's lab will be available to the project.

%--------------------------------------------------------------------------------------------%
%xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx%
%--------------------------------------------------------------------------------------------%

\newpage
\pagenumbering{gobble}

\begin{center}
\Large{\textbf{\\Data Management Plan}}
\end{center}

This data management plan covers all of the data products to be produced in this project. It includes all source code that implements the ecosystem of GPU resource management, any source code added to LITMUS$^{\textrm{RT}}$,  the suite of case-study programs, and scripts used to test and evaluate the implementation of the proposed components, along with the data collected during evaluation efforts. 

\paragraph{Source code.} All source code will be freely available for direct download from public web servers maintained by the UTD Computer Science Department under an open source GPL license.  Additional case-study test programs and scripts will be freely available under the open source BSD license and directly downloadable from public web servers maintained by the UTD Computer Science Department. Our intend in making these data products freely available is that other researchers can verify evaluation results, and relevant research conducted by other interested groups can be facilitated.

\paragraph{Evaluation data.} Evaluation data (including raw data and resulting graphs) can be several terabytes in size. It is thus currently not feasible to make this data available to the public through direct download. We will store and back up such data on secured servers maintained by the UT-Dallas Computer Science Department. Interested parties may request copies of this data through making arrangement for private direct download or using the postal system.

\paragraph{Privacy.} No personal data will be acquired in this project. There are no privacy concerns for the storage of data.

\paragraph{Long-term storage.} All data products produced in this project will also be archived in the UTD computer science computing repository for permanent storage.

%\end{comment}

%--------------------------------------------------------------------------------------------%
%xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx%
%--------------------------------------------------------------------------------------------%

\end{document}
